---
title: "Activity: Bird Species"
author: "Alvin Chen"
date: "Wednesday, June 7, 2023"
output: html_notebook
---


# Set Up

```{r}
# Load `tidyverse` and `dcData` packages

library(tidyverse)
library(dcData)

# Load the `OrdwayBirds` data set into your RStudio environment from the `dcData` package

data("OrdwayBirds", package = "dcData")

# Show a few commands used to inspect the data set a few different ways

?OrdwayBirds
head(OrdwayBirds)
view(OrdwayBirds)

```

# Activity

When you're finished with this activity, you will have a graph that shows what time of year various species appear at the Katherine Ordway Natural History Study Area in Inver Grove Heights, MN.

### Step 0

Before we begin, the book gives us some instructions to select a few key variables and clean up the date formatting.  Use the commands given in the book, but be sure to study the functions used so you know what is happening.  In this case, we select `SpeciesName`, `Month`, and `Day`. Then a `mutate()` command converts each variable to character, and then converts the resulting "character string" to a number.

```{r}
# Get the data table & clean up dates (see Data Computing eBook)

OrdwayBirds1 <-
  OrdwayBirds %>%
  select( SpeciesName, Month, Day ) %>%
  mutate( Month = as.numeric(as.character(Month)), 
          Day = as.numeric(as.character(Day)))

# Inspect resulting data

head(OrdwayBirds)
view(OrdwayBirds)

```

Observations: Upon inspecting the data set I can see that there are a bunch of bird secies and a lot od data concerning each individual bird citing. 

### Step 1


```{r}
# unique species in original data (including mis-spellings)

unique_species_org <-
  n_distinct(OrdwayBirds$SpeciesName) #275 unique species

# unique species in the clean list

unique_species_clean <-
  n_distinct(OrdwayBirds1$SpeciesName) #275 unique species

```

Observation: I noticed that the amount od species stayed the same but the variables went from 26 to 3.

### Step 2


```{r}
# inspect OrdwaySpeciesNames

head(OrdwaySpeciesNames)
view(OrdwaySpeciesNames)

# inner join

joined_birds <-
  OrdwayBirds %>%
  inner_join(OrdwaySpeciesNames)
  
# inspect the data after the join

head(joined_birds)
view(joined_birds)

```

Which variable(s) was used for matching cases?
    -SpeciesName Variable

What were the variable(s) that will be added?
    -SpeciesNameCleaned variable will be added 
    
Observation: the OrdwaySPeciesNames table cleaned up the names ans compressed so they can be categorized easier.

### Step 3

We want to identify the top 5-6 bird species sighted at the Katherine Ordway Natural History Study Area.

```{r}
# count sightings by species in descending order

sightings_species <- OrdwayBirds %>%
  count(SpeciesName, sort = TRUE) 

# a hint in the text book recommends displaying the top 10 species to choose our threshold

sightings_species <- OrdwayBirds %>%
  count(SpeciesName, sort = TRUE) %>%
  head(10)

# define a sightings threshold; produce a table with only birds belonging to major species

major_bird <-
  OrdwayBirds %>%
  filter(SpeciesName %in% sightings_species$SpeciesName)


```

Observations: the new table is much more orginized and setting a threshold allows us to focus in on specific species.

### Step 4

We now want to produce a month-by-month count for each of the major species, and then construct a chart to tell the story of what time of year the various species appear.

```{r}
# prepare glyph-ready data table called `ByMonth`

ByMonth <-
  major_bird %>%
  group_by(SpeciesName, Month) %>%
  count()


```

Now use the ggplot template below to generate the desired graph. (Remark you can also use esquisser or mplot to create a line chart instead of the barchart)



```{r}
# ggplot commands

ggplot(ByMonth) +
  aes(x = Month, y = n, colour = SpeciesName) +
  geom_point(shape = "circle", size = 4L) +
  scale_color_hue(direction = 1) +
  theme_linedraw()

```

According to the graph, we can answer the questions posed:

##### 1. Which species are present year-round?  

-American Goldfinch 
-Black-capped Chickadee
-Blue Jay
-Slate-colored Junco

##### 2. Which species are migratory, that is, primarily present in one or two seasons?

-Catbird
-Field Sparrow
-House Wren
-Robin
-Song Sparrow
-Tree Sparrow


##### 3. What is the peak month for each major species? (bullet list)

-	American Goldfinch: October 
- Black-capped Chickadee: November 
- Blue Jay: May
- Catbird: May
- Field Sparrow: May
- House Wren: July
- Robin: May
- Slate-colored Junco: October
- Song Sparrow: April
- Tree Sparrow: March

##### 4. Which major species are seen in good numbers for at least 6 months of the year?

-American Goldfinch
-Black-capped Chickadee
-Tree swallow


Observations: After completing step four I can observe that allows for people to accuratley determine when its best to go to see each species.
```{r}

```

